Milk intake and stroke mortality in the Japan Collaborative Cohort Study - a Bayesian survival analysis

### 2020-08-26 Paperfolder/MDPI_submission/Revision01/MDPI_submission/Paperfolder.tex
updated ready to send to Yatsuya

% Abstract (Do not insert blank lines, i.e. \\)
\abstract{The aim was to further examine the relationship between milk
intake and stroke mortality among the Japanese population. We used data
from the Japan Collaborative Cohort (JACC) Study (total number of
participants = 110585, age range: 40-79) to estimate the posterior
acceleration factors (AF) as well as the hazard ratios (HR) comparing
individuals with different milk intake frequencies against those who
never consumed milk at the study baseline. These estimations were
computed through a series of Bayesian survival models that employed a
Markov Chain Monte Carlo simulation process. 100000 posterior samples
were generated separately through four independent chains after model
convergency were confirmed. Posterior probabilites that daily milk
consumers had lower hazard or delayed mortality from strokes compared to
non-consumers was 99.0\% and 78.0\% for men and women, respectively.
Accordingly, the estimated posterior means of AF and HR for daily milk
consumers were 0.88 (95\% Credible Interval, CrI: 0.81, 0.96) and 0.80
(95\% CrI: 0.69, 0.93) for men and 0.97 (95\% CrI: 0.88, 1.10) and 0.95
(95\% CrI: 0.80, 1.17) for women. In conclusion, data from the JACC
study has provided strong evidence that daily milk intake among Japanese
men was associated with delayed and lower hazard of mortality from
stroke especially cerebral infarction.}

@@

Eastern Asian populations were reported to have higher burden from
either mortality or morbidity from stroke than populations in European
or American regions \citep{Kim2014}. Dairy food, especially milk has
been suggested to reduce stroke risk by nearly 7\% for each 200 g
increment of daily consumption \citep{DeGoede2016}. Although 2 daily
servings of milk or dairy products is recommended in Japan
\citep{Yoshiike2007}, the actual per capita intake (\(\approx\) 63
g/day) of these food groups is much lower and less frequent than that in
Western countries \citep{Saito2019}. Given that previous reports have
also indicated no significant
\citep{Iso1999, sauvaget2003intake, Elwood2004} or even positive
associations \citep{Larsson2009}, it would be of interest to provide
information on the association within a context that most individuals
have much lower range of milk intake compared to most of the previous
studies. Whether people with such a low level intake of milk can still
benefit against stroke would require elucidation.

Moreover, a more intuitive interpretation would be available if we were
able to show the probabilities and how certain the existing data can
provide evidence about whether drinking milk can delay or lower the
hazard of dying from stroke. This would require transformation of the
research questions into ``Do individuals with milk intake habit have
lower hazard of dying from a stroke event compared with people who never
consume milk?'' Another way of asking the same research question would
be ``Can a event of dying from stroke be postponed by having a milk
intake habit?'' The later question is indeed directly comparing the
length of times before events or the speed from the same departure point
(entry of the study) to a fatal stroke event. Accelerated failure time
models (AFT) under Bayesian framework are convenient tools that would
help avoid worrying about the assumptions of repeatability of a large
scale prospective study, and can calculate the exact proportion of
posterior estimates (a hazard ratio) that is lower than 1. AFT model is
helpful in estimating acceleration factors which can be interpreted as
the speed for developing an event of interest. This speed parameter
shows how faster/slower individuals in one exposure group might have an
event compared to others among different exposure groups
\citep{Wei1992, ibrahim2014b}.

@@

Frequency of milk intake during the preceding year of the baseline was
assessed by FFQ from ``never'', ``1-2 times/month'', ``1-2 times/week'',
``3-4 times/week'', and ``Almost daily''. The exact amount of milk
consumption was difficult to assess here. However, good reproducibility
and validity were confirmed previously (Spearman rank correlation
coefficient between milk intake frequency and weighed dietary record for
12 days was 0.65) \citep{Date2005}. From the same validation study
\citep{Date2005} on the FFQ, daily consumers of milk was found to have a
median intake of 146 gram per day. In the baseline year of the JACC
study (between 1988 and 1990), most ``milk and dairy product''
consumption (92.1\%) was in the form of whole milk \citep{eigyogenjyo}.

@@

Analyses were stratified by sex as difference in milk consumption levels
and mortality rate was suggested previously
\citep{wang_milk_2015, Talaei_2016}. We calculated sex-specific means
(standard deviation, sd) and proportion of selected baseline
characteristics according to the frequency of milk intake. Age-adjusted
stroke mortality rate were expressed as per 1000 person-year predicted
through poisson regression models.

Full parametric proportional hazard models under Bayesian framework with
Weibull distribution were fitted using Just Another Gibbs Sampler (JAGS)
program \citep{Plummer2003} version 4.3.0 in R version 4.0.1
\citep{RCT2020}. JAGS program is similar to the OpenBUGS
\citep{Lunn2009} project that uses a Gibbs sampling engine for MCMC
simulation. In the current analysis, we specified non-informative prior
distributions for each of the parameters in our models
(\(\beta_n \sim N(0, 1000)\), and
\(\kappa_{\text{shape}} \sim \Gamma(0.001, 0.001)\)). The
Brooks-Gelman-Rubin diagnostic \citep{Brooks1998} was used to refine the
approximate point of convergence, the point when the ratio of the chains
is stable around 1 and the within and between chain variability start to
reach stability was visually checked. The auto-correlation tool further
identified if convergence has been achieved or if a high degree of
auto-correlation exists in the sample. Then, the number of iterations
discarded as `burn-in' was chosen. All models had a posterior sample
size of 100000 from four separated chains with a ``burn-in'' of 2500
iterations. Posterior means (sd) and 95\% Credible Intervals (CrI) of
the estimated hazard ratios (HRs) as well as acceleration factors (AFs)
were presented for each category of milk intake frequency taking the
``never'' category as the reference. Posterior probabilities that the
estimated hazard of dying from stroke for the milk intake for frequency
that higher or equal to ``1-2 times/month'' is smaller compared with
those who chose ``never'' to their milk intake frequency were calculated
as \(Pr(\text{HR} < 1)\).

@@

@@

The total follow-up was 1555073 person-years (median = 19.3 years),
during which 2675 deaths from stroke were confirmed (1352 men and 1323
women). Among these stroke mortality, 952 were hemorrhagic stroke (432
men and 520 women), and 957 were cerebral infarction (520 men and 437
women). Number of deaths from causes other than stroke were 18868 (10731
men and 8137 women); 5493 (2022 men and 3471 women) dropped out from the
follow-up (5.8\%); 67349 (25281 men and 42068 women) were censored at
the end of follow-up (71.4\%). The median (interquartile range, IQR) of
follow-up years for hemorrhagic stroke mortality were 9.9 (5.4, 12.3) in
men and 10.9 (5.9, 15.1) in women; the median (IQR) of follow-up years
for cerebral infarction mortality were 11.2 (7.1, 15.3) and 11.8 (7.9,
16.4) in men and women. Age-adjusted stroke mortality rates for each
category of milk intake frequency was estimated to be 1.8, 2.0, 1.7,
1.6, 1.5 per 1000 person-year for men and 1.3, 1.4, 1.2, 1.1, 1.2 per
1000 person-year for women.

@@

Detailed results from the Bayesian survival models (crude, age-adjusted
and multivariable-adjusted) according to the frequency of milk intake
separated by sex are listed in \textbf{Table \ref{tab:tab2}} (men) and
\textbf{Table \ref{tab:tab3}} (women). Compared to those who never had
milk, both men and women had slower speed and lower hazard of dying from
total stroke in crude models. Velocities that milk consumers dying from
stroke is slower by a crude acceleration factor (AF) between 0.79 (sd =
0.05; 95\% CrI: 0.74, 0.90) and 0.93 (sd = 0.04; 95\% CrI: 0.85, 1.02)
compared with non-consumers. Chances that the posterior crude HRs were
estimated to be lower than 1 for those who had at least 1-2 times/month
was higher than 86.5\% in men and greater than 94.6\% in women. However,
lower hazard and delayed time-to-event was observed to remain after age
or multivariable adjustment only among daily male

@@

\noindent milk consumers. Specifically, the mean (sd; 95\% CrI) of
posterior multivariable-adjusted AF and HR for daily male consumers of
milk were 0.88 (sd = 0.05; 95\% CrI: 0.81, 0.96) and 0.80 (sd = 0.07;
95\% CrI: 0.69, 0.93) with a probability of 99.0\% to be smaller than
the null value (=1). Daily female milk consumers had posterior AFs and
HRs that were distributed with means of 0.97 (sd = 0.09; 95\% CrI: 0.88,
1.10) and 0.95 (sd = 0.12; 95\% CrI: 0.80, 1.17) which had about 78.0\%
of chance that their HRs could be smaller than 1.

@@

These findings showed similar negative effect estimates reported
previously \citep{wang_milk_2015} using data from the same cohort, in
which the outcomes of interests were focused on cardiovascular diseases
and all-cause mortality. Moreover, we have further updated the results
with more comprehensive and straightforward evidence about whether and
how certain the data had shown about daily consumption of milk is
contributing to a postponed stroke (mostly cerebral infarction)
mortality event among Japanese men. A recent dose-response meta-analysis
of 18 prospective cohort studies had shown a similar negative
association \citep{DeGoede2016} between milk consumption and risk of
stroke. The same meta-analysis also reported a greater reduction of risk
of stroke (18\%) for East Asian population in contrast with the 7\% less
risk in the pooled overall finding for all populations combined.
Benefits of increased milk intake might be particularly noticeable in
East Asian countries where strokes are relatively more common, and milk
consumption is much lower than those studies conducted among European or
American populations \citep{dehghan2018association}. However, the Life
Span Study \citep{sauvaget2003intake}, which was conducted about 10
years earlier than the JACC Study, reported null association between
milk intake and fatal stroke among men and women combined who survived
from the Hiroshima and Nagasaki atomic bomb. The difference could be
largely explained by the different targeting populations between the two
studies. \citet{kondo2013consumption} also reported null association in
either men or women from the NIPPON DATA80 study. In fact, the exact
number of stroke mortality events was about 85\% less and the number of
participants in the NIPPON DATA80 was about 90\% less than those in the
JACC study database, their null association might likely due to its
limited statistical power. \citet{umesawa2008dietary} reported that
dairy calcium intake was inversely associated with ischemic stroke
mortality risk which could be considered as one of the potential
evidence that supported our findings. Furthermore, stronger inverse
association between milk consumption and stroke mortality in men rather
than in women was found in the Singapore Chinese Health Study
\citep{Talaei_2016} as well. The reasons for this gender differences is
currently unknown. It might be contributed by generally better/healthier
lifestyles (much less smokers and alcohol drinkers) in women regardless
of milk intake frequency, or maybe due to other factors that was not
available in/considered by our models such as intake of calcium
supplements. The probably existing beneficial effect by milk intake is
probably less evident in women than that in men. Further investigation
is needed.

@@

\noindent
\begin{tabular}{@{}ll}
JACC & Japan Collaborative Cohort \\
FFQ & Food Frequency Questionnaire \\
MCMC & Markov Chain Monte Carlo \\
JAGS & Just Another Gibbs Samplers \\
AFT & accelerated failure time \\
HR & hazard ratio \\
AF & acceleration factor \\
sd & standard deviation \\
CrI & credible interval \\
\end{tabular}}

>>>>

% Abstract (Do not insert blank lines, i.e. \\)
\abstract{The aim was to further examine the relationship between milk
intake and stroke mortality among the Japanese population. We used data
from the Japan Collaborative Cohort (JACC) Study (total number of
participants = 110585, age range: 40-79) to estimate the posterior
acceleration factors (AF) as well as the hazard ratios (HR) comparing
individuals with different milk intake frequencies against those who
never consumed milk at the study baseline. These estimations were
computed through a series of Bayesian survival models that employed a
Markov Chain Monte Carlo simulation process. 100000 posterior samples
were generated separately through four independent chains after model
convergency was confirmed. Posterior probabilites that daily milk
consumers had lower hazard or delayed mortality from strokes compared to
non-consumers was 99.0\% and 78.0\% for men and women, respectively.
Accordingly, the estimated posterior means of AF and HR for daily milk
consumers were 0.88 (95\% Credible Interval, CrI: 0.81, 0.96) and 0.80
(95\% CrI: 0.69, 0.93) for men and 0.97 (95\% CrI: 0.88, 1.10) and 0.95
(95\% CrI: 0.80, 1.17) for women. In conclusion, data from the JACC
study has provided strong evidence that daily milk intake among Japanese
men was associated with delayed and lower hazard of mortality from
stroke especially cerebral infarction.}

@@

Eastern Asian populations were reported to have higher burden from both
mortality and morbidity from stroke than populations in European or
American regions \citep{Kim2014}. Dairy foods, especially milk has been
suggested to related with decreased stroke risk by nearly 7\% for each
200 g increment of daily consumption \citep{DeGoede2016}. Although 2
daily servings of milk or dairy products is recommended in Japan
\citep{Yoshiike2007}, the actual per capita intake (\(\approx\) 63
g/day) of these food groups is much lower and less frequent than that in
Western countries \citep{Saito2019}. Given that previous reports have
also indicated no significant
\citep{Iso1999, sauvaget2003intake, Elwood2004} or even positive
associations \citep{Larsson2009}, it would be of interest to provide
information on the association within a context that most individuals
have much lower range of milk intake compared to most of the previous
studies. Whether people with such a low level intake of milk can still
benefit against stroke would require elucidation.

Moreover, a more intuitive or straightforward interpretation would be
available if we were able to show the probabilities and the certainty
that the existing data can provide evidence about whether drinking milk
can delay or lower the hazard of dying from stroke. This would require
transformation of the research question into ``Do individuals with milk
intake habit have lower hazard of dying from a stroke event?'' Another
way of asking the same research question would be ``Can a event of dying
from stroke be postponed by having a milk intake habit?'' The later
question is indeed directly comparing the length of times before events
or the speed from the same departure point (entry of the study) to a
fatal stroke event. Accelerated failure time models (AFT) under Bayesian
framework are convenient tools that would help avoid worrying about the
assumptions of repeatability of a large scale prospective study, and can
calculate the exact proportion of posterior estimates (hazard ratios,
HR) that is lower than 1. AFT model is helpful in estimating
acceleration factors (AF) which can be interpreted as the velocity for
developing an event of interest. This velocity parameter shows how
faster/slower individuals in one exposure group might have an event
compared to others in different exposure groups
\citep{Wei1992, ibrahim2014b}.

@@

Frequency of milk intake during the preceding year of the baseline was
assessed by FFQ as ``never'', ``1-2 times/month'', ``1-2 times/week'',
``3-4 times/week'', and ``Almost daily''. The exact amount of milk
consumption was difficult to assess here. However, good reproducibility
and validity were confirmed previously (Spearman rank correlation
coefficient between milk intake frequency and weighed dietary record for
12 days was 0.65) \citep{Date2005}. From the same validation study
\citep{Date2005} on the FFQ, daily consumers of milk was found to have a
median intake of 146 gram per day. At the baseline of the JACC study
(between 1988 and 1990), most ``milk and dairy product'' consumption
(92.1\%) was in the form of whole milk \citep{eigyogenjyo}.

@@

Analyses were stratified by sex as difference in milk consumption levels
and mortality rate was suggested previously
\citep{wang_milk_2015, Talaei_2016}. We calculated sex-specific means
(standard deviation, SD) and proportion of selected baseline
characteristics according to the frequency of milk intake. Age-adjusted
stroke mortality rate were expressed as per 1000 person-year predicted
through poisson regression models.

Full parametric proportional hazard models under Bayesian framework with
Weibull distribution were fitted using Just Another Gibbs Sampler (JAGS)
program \citep{Plummer2003} version 4.3.0 in R version 4.0.1
\citep{RCT2020}. JAGS program is similar to the OpenBUGS
\citep{Lunn2009} project that uses a Gibbs sampling engine for MCMC
simulation. In the current analysis, we specified non-informative prior
distributions for each of the parameters in our models
(\(\beta_n \sim N(0, 1000)\), and
\(\kappa_{\text{shape}} \sim \Gamma(0.001, 0.001)\)). The
Brooks-Gelman-Rubin diagnostic \citep{Brooks1998} was used to refine the
approximate point of convergence, the point when the ratio of the chains
is stable around 1 and the within and between chain variability start to
reach stability was visually checked. The auto-correlation tool further
identified if convergence has been achieved or if a high degree of
auto-correlation exists in the sample. Then, the number of iterations
discarded as `burn-in' was chosen. All models had a posterior sample
size of 100000 from four separated chains with a ``burn-in'' of 2500
iterations. Posterior means (SD) and 95\% Credible Intervals (CrI) of
the estimated hazard ratios (HRs) as well as acceleration factors (AFs)
were presented for each category of milk intake frequency taking the
``never'' category as the reference. Posterior probabilities that the
estimated hazard of dying from stroke for the milk intake for frequency
that higher or equal to ``1-2 times/month'' is smaller compared with
those who chose ``never'' to their milk intake frequency were calculated
as \(Pr(\text{HR} < 1)\).

@@

\multicolumn{6}{l}{Abbreviations: JACC: Japan Collaborative Cohort; SD: standard deviation; BMI: body mass index.   }\\

@@

The total follow-up was 1555073 person-years (median = 19.3 years),
during which 2675 deaths from stroke were confirmed (1352 men and 1323
women). Among these stroke mortality, 952 were hemorrhagic stroke (432
men and 520 women), and 957 were cerebral infarction (520 men and 437
women). Number of deaths from causes other than stroke was 18868 (10731
men and 8137 women); 5493 (2022 men and 3471 women) dropped out from the
follow-up (5.8\%); 67349 (25281 men and 42068 women) were censored at
the end of follow-up (71.4\%). The median (interquartile range, IQR) of
follow-up years for hemorrhagic stroke mortality were 9.9 (5.4, 12.3) in
men and 10.9 (5.9, 15.1) in women; the medians (IQR) of follow-up years
for cerebral infarction mortality were 11.2 (7.1, 15.3) and 11.8 (7.9,
16.4) in men and women, respectively. Age-adjusted stroke mortality
rates for each category of milk intake frequency was estimated to be
1.8, 2.0, 1.7, 1.6, 1.5 per 1000 person-year for men and 1.3, 1.4, 1.2,
1.1, 1.2 per 1000 person-year for women.

@@

Detailed results from the Bayesian survival models (crude, age-adjusted
and multivariable-adjusted) according to the frequency of milk intake
separated by sex are listed in \textbf{Table \ref{tab:tab2}} (men) and
\textbf{Table \ref{tab:tab3}} (women). Compared to those who never had
milk, both men and women had slower speed and lower hazard of dying from
total stroke in crude models. Velocities that milk consumers dying from
stroke is slower by a crude acceleration factor (AF) between 0.79 (SD =
0.05; 95\% CrI: 0.74, 0.90) and 0.93 (SD = 0.04; 95\% CrI: 0.85, 1.02)
compared with non-consumers. Chances that the posterior crude HRs were
estimated to be lower than 1 for those who had at least 1-2 times/month

@@

\noindent was higher than 86.5\% in men and greater than 94.6\% in
women. However, lower hazard and delayed time-to-event was observed to
remain after age or multivariable adjustment only among daily male milk
consumers. Specifically, the mean (SD; 95\% CrI) of posterior
multivariable-adjusted AF and HR for daily male consumers of milk were
0.88 (SD = 0.05; 95\% CrI: 0.81, 0.96) and 0.80 (SD = 0.07; 95\% CrI:
0.69, 0.93) with a probability of 99.0\% to be smaller than the null
value (=1). Daily female milk consumers had posterior AFs and HRs that
were distributed with means of 0.97 (SD = 0.09; 95\% CrI: 0.88, 1.10)
and 0.95 (SD = 0.12; 95\% CrI: 0.80, 1.17) which had about 78.0\% of
chance that their HRs could be smaller than 1.

@@

These findings showed similar negative effect estimates that were
reported previously \citep{wang_milk_2015} using data from the same
cohort, in which the outcomes of interests were focused on
cardiovascular diseases and all-cause mortality. Moreover, we have
further updated the results with more comprehensive and straightforward
evidence about whether and how certain the data had shown about daily
consumption of milk is contributing to a postponed stroke (mostly
cerebral infarction) mortality event among Japanese men. A recent
dose-response meta-analysis of 18 prospective cohort studies had shown a
similar negative association \citep{DeGoede2016} between milk
consumption and risk of stroke. The same meta-analysis also reported a
greater reduction of risk of stroke (18\%) for East Asian population in
contrast with the 7\% less risk in the pooled overall finding for all
populations combined. Benefits of increased milk intake might be
particularly noticeable in East Asian countries where strokes are
relatively more common, and milk consumption is much lower than those
studies conducted among European or American populations
\citep{dehghan2018association}. However, the Life Span Study
\citep{sauvaget2003intake}, which was conducted about 10 years earlier
than the JACC Study, reported null association between milk intake and
fatal stroke among men and women combined who survived from the
Hiroshima and Nagasaki atomic bomb. The difference could be largely
explained by the different targeting populations between the two
studies. \citet{kondo2013consumption} also reported null association in
either men or women from the NIPPON DATA80 study. In fact, the exact
number of stroke mortality events was about 85\% less and the number of
participants in the NIPPON DATA80 was about 90\% less than those in the
JACC study database, their null association might likely due to its
limited statistical power. \citet{umesawa2008dietary} reported that
dairy calcium intake was inversely associated with ischemic stroke
mortality risk which could be considered as one of the potential
evidence that supported our findings. Furthermore, stronger inverse
association between milk consumption and stroke mortality in men rather
than in women was found in the Singapore Chinese Health Study
\citep{Talaei_2016} as well. The reasons for this gender differences is
currently unknown. It might be contributed by generally better/healthier
lifestyles (such as much less smokers and alcohol drinkers) in women
regardless of milk intake frequency, or maybe due to other factors that
was not available in/considered by our models such as intake of calcium
or other supplements. The probably existing beneficial effect by milk
intake is probably less evident in women than that in men. Further
investigation is needed.

@@

\noindent
\begin{tabular}{@{}ll}
JACC & Japan Collaborative Cohort \\
FFQ & Food Frequency Questionnaire \\
MCMC & Markov Chain Monte Carlo \\
JAGS & Just Another Gibbs Samplers \\
AFT & accelerated failure time \\
HR & hazard ratio \\
AF & acceleration factor \\
SD & standard deviation \\
CrI & credible interval \\
\end{tabular}}
